A Novel Solution to Resonant Scattering with Self-Consistent Boundary Conditions

### 2022-04-06 ms.tex
First pass at addressing reviewer comments

 ``Hummer Case II-b"  \citep{1962MNRAS.125...21H} will be used for the redistribution function, for which the incoming photon is absorbed by the atom according to the natural broadening profile in the rest frame, then is re-emitted with a dipole phase function $g(\vec{n},\vec{n}^\prime)=(3/16\pi)(1+[\vec{n}\cdot \vec{n}^\prime]^2)$, which is appropriate for a 1s-2p transition \citep{1982qe}, and then is averaged over a Maxwell-Boltzmann distribution of speeds for the atom. The result can be written
\be
j_{\rm sc}(\vec{x},\vec{n},\nu) & = & k \int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\, 
\int d\Omega^\prime \int d\nu^\prime \,
g(\vec{n},\vec{n}^\prime) 
\nonumber \\ & \times & 
\delta \left( \nu - \nu^\prime - \nu_0 \vec{v} \cdot (\vec{n}-\vec{n}^\prime)/c \right)
\left( \frac{\Gamma/4\pi^2}{ \left(\nu^\prime - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}^\prime/c \right)^2 + (\Gamma/4\pi)^2 } \right)  \,
I(\vec{x},\vec{n}^\prime,\nu^\prime)
\nonumber \\ & = & 4\pi k \int d\Omega^\prime \int d\nu^\prime R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) I(\vec{x},\vec{n}^\prime,\nu^\prime),
\ee
which defines the Case II-b redistribution function
\be
R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) & = & \frac{ g(\vec{n},\vec{n}^\prime) }{ 4\pi }
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\delta \left( \nu - \nu^\prime - \nu_0 \vec{v} \cdot (\vec{n}-\vec{n}^\prime)/c \right)
\left( \frac{\Gamma/4\pi^2}{ \left(\nu^\prime - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}^\prime/c \right)^2 + (\Gamma/4\pi)^2 } \right)
\ee
found in \citet{1962MNRAS.125...21H}.

The integral of the redistribution function over outgoing and incoming frequency are
\be
\int d\nu\ R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
& = & \frac{1}{4\pi} g(\vec{n},\vec{n}^\prime) \phi(\nu^\prime)
\ee 
and
\be
\int d\nu^\prime \ R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
& = & \frac{1}{4\pi} g(\vec{n},\vec{n}^\prime) \phi(\nu)
\ee 
where the right hand side is the usual Voigt function, the thermal average of the Lorentzian. The former result implies that the integrated source and sink terms for scattering cancel for $p=1$. In addition, $4\pi R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime)/\phi(\nu^\prime) $ is the normalized distribution for the outgoing $\vec{n}$ and $\nu$ given the incoming $\vec{n}^\prime$ and $\nu^\prime$.

This probability distribution can be used to define the moments of the frequency shift
\be
\langle \delta \nu^n \rangle & = & \frac{ \int d\nu^\prime (\nu^\prime-\nu)^n R}{\int d\nu^\prime R}
= \frac{1}{\phi(\nu)}
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\left( \frac{\nu_0 \vec{v} \cdot (\vec{n}^\prime-\vec{n}) }{c} \right)^n
\left( \frac{\Gamma/4\pi^2}{ \left(\nu - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}/c \right)^2 + (\Gamma/4\pi)^2 } \right),
\ee
which are functions of $\nu$, $\vec{n}$ and $\vec{n}^\prime$. These integrals can be evaluated in terms of the dimensionless moments of the parallel velocity distribution, defined as
\be
\langle u_\parallel^n \rangle(x,a) & = & \frac{a/\pi }{\mathcal{H}(x,a)} \int 
\frac{du_\parallel u_\parallel^n e^{-u_\parallel^2}  }{(x-u_\parallel)^2 + a^2}.
\ee
The end results for the first and second moments are
\be
\langle \delta \nu \rangle & = & -\langle u_\parallel \rangle \left( 1 - \vec{n} \cdot \vec{n}^\prime \right) \Delta 
\\
\langle \delta \nu^2 \rangle & = & 
\left[ \langle u_\parallel^2 \rangle
\left( 1 - \vec{n} \cdot \vec{n}^\prime \right)^2
+ \frac{1}{2} \left( 1 - \left( \vec{n} \cdot \vec{n}^\prime\right)^2 \right) \right] \Delta^2 .
\ee

For small frequency shifts $\nu^\prime-\nu$, the incoming intensity may be expanded as 
\be
I(\vec{x},\vec{n}^\prime,\nu^\prime) & \simeq  &
I(\vec{x},\vec{n}^\prime,\nu) 
+ 
\frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }(\nu^\prime-\nu)
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2} (\nu^\prime-\nu)^2 + ...
\ee 
and the Fokker-Planck expansion of $j_{\rm sc}$ is
\be
j_{\rm sc}(\vec{x},\vec{n},\nu) & \simeq  & 4\pi k \int d\Omega^\prime \int d\nu^\prime R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
\left[ 
I(\vec{x},\vec{n}^\prime,\nu) + \frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }(\nu^\prime-\nu)
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2 }(\nu^\prime-\nu)^2
\right]
\nonumber \\ 
& =& 
k\phi(\nu) \int d\Omega^\prime g 
\left[ 
I(\vec{x},\vec{n}^\prime,\nu) 
+ \frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }
\langle \delta \nu \rangle
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2 }
\langle \delta \nu^2 \rangle
\right].
\ee 
To perform the angular integrals, the Eddington approximation for the angular dependence is inserted with the following result
\be
j_{\rm sc} & = & k\phi J - k\phi \Delta \langle u_\parallel \rangle \left( \frac{\partial J}{\partial \nu} - \frac{6}{5} \vec{n} \cdot \frac{\partial \vec{H}}{\partial \nu} \right)
+ \frac{1}{2} \Delta^2 k\phi \left[ 
\frac{\partial^2 J}{\partial \nu^2} \left( \frac{7}{5} \langle u_\parallel^2 \rangle + \frac{3}{10} \right)
- \frac{12}{5} \langle u_\parallel^2 \rangle 
\vec{n} \cdot \frac{\partial^2 \vec{H}}{\partial \nu^2} \right]
\nonumber \\ & \simeq & 
k\phi J - k\phi \Delta \langle u_\parallel \rangle  \frac{\partial J}{\partial \nu} 
+ \frac{1}{2} \Delta^2 k\phi \left( \frac{7}{5} \langle u_\parallel^2 \rangle + \frac{3}{10} \right)
\frac{\partial^2 J}{\partial \nu^2} .
\label{eq:jsc}
\ee
The first term in Equation (\ref{eq:jsc}), $k\phi J$, represents re-emission of the photon through de-excitation of the atom. It cancels the $-k\phi J$ term in Equation (\ref{eq:rteqn}) that corresponds to excitation of the atom. The terms involving frequency derivatives of $\vec{H}$, if carried through the calculation, end up giving terms smaller than the largest terms by a factor of $1/x^2$, which is small in the line wing. These terms are ignored from here onward.

@@

\right) = 0,

@@

Near the line center, there are two independent solutions which can be written
\be \label{eq:wkb_J}
J(n,\sigma,s) & = & a_1 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/2} J_{-3/4}\left( k_c \left[ \frac{\sigma}{\sigma_{\rm tp} }\right]^{2/3} \right)
\nonumber \\ &  + &  a_2 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/2} J_{3/4}\left( k_c \left[ \frac{\sigma}{\sigma_{\rm tp} }\right]^{2/3} \right),
\ee
where $J_\alpha(x)$ are Bessel functions of the first kind, $a_1$ and $a_2$ are normalization constants, $\sigma_{tp}$ is given in Equation (\ref{eq:sigma_tp}), and $k_c=3\kappa_n \Delta \sigma_{\rm tp}/2k$. The solution far from line center is 
\be \label{eq:wkb_J_far}
J(n,\sigma,s) & = & a_3 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/6} Ai\left(k_c^{2/3} \left[ \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{2/3} - 1 \right]\right),
\ee
where $Ai(x)$ is the Airy function of the first kind. Again, $a_3$ is a normalization constant. We can enforce the discontinuity in Equation (\ref{eq:matching_condition_2}) to find 
\be \label{eq:a2}
a_2 & = & - \frac{\Gamma(7/4) \sigma_{\rm tp}}{ 2^{1/4} k_c^{3/4}} \frac{6^{1/2}}{8} n^2 \frac{E}{kR^3}.
\ee
The discontinuity sets the value of $a_2$, and the corresponding term in Equation (\ref{eq:wkb_J}) is non-resonant. The discontinuity does not set the value of $a_1$ as that solution does not have a term linear in $\sigma$. Matched asymptotic expansions \citep{1999amms.book.....B} allows us to relate the normalization coefficients $a_1$ and $a_3$. The matching conditions are
\be \label{eq:a1/a2}
\frac{a_1}{a_2} & = & \frac{ \sin \left( \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} + \frac{\pi}{8} \right) }{ \sin \left(  \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} - \frac{\pi}{8} \right) }
\ee
\be \label{eq:a3/a2}
\frac{a_3}{a_2} & = & \frac{ k_c^{-1/3} }{ \sin \left(  \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} - \frac{\pi}{8} \right) }.
\ee
By setting the to denominator of this expression equal to zero and solving for the eigenfrequency $\gamma_{nm}$ contained in $\sigma_{\rm tp}$, we find the dispersion relation in Equation (\ref{eq:gamma_nm}).

>>>>

 ``Hummer Case II-b"  \citep{1962MNRAS.125...21H} will be used for the redistribution function, for which the incoming photon is absorbed by the atom according to the natural broadening profile in the rest frame, then is re-emitted with a dipole phase function $g(\vec{n},\vec{n}^\prime)=(3/16\pi)(1+[\vec{n}\cdot \vec{n}^\prime]^2)$, which is appropriate for a 1s-2p transition \citep{1982qe}, and then is averaged over a Maxwell-Boltzmann distribution of speeds for the atom. The result can be written
\be
j_{\rm sc}(\vec{x},\vec{n},\nu) & = & k \int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\, 
\int d\Omega^\prime \int d\nu^\prime \,
g(\vec{n},\vec{n}^\prime) 
\nonumber \\ & \times & 
\delta \left( \nu - \nu^\prime - \nu_0 \vec{v} \cdot (\vec{n}-\vec{n}^\prime)/c \right)
\left( \frac{\Gamma/4\pi^2}{ \left(\nu^\prime - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}^\prime/c \right)^2 + (\Gamma/4\pi)^2 } \right)  \,
I(\vec{x},\vec{n}^\prime,\nu^\prime)
\nonumber \\ & = & 4\pi k \int d\Omega^\prime \int d\nu^\prime R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) I(\vec{x},\vec{n}^\prime,\nu^\prime),
\ee
which defines the Case II-b redistribution function
\be
R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) & = & \frac{ g(\vec{n},\vec{n}^\prime) }{ 4\pi }
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\delta \left( \nu - \nu^\prime - \nu_0 \vec{v} \cdot (\vec{n}-\vec{n}^\prime)/c \right)
\left( \frac{\Gamma/4\pi^2}{ \left(\nu^\prime - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}^\prime/c \right)^2 + (\Gamma/4\pi)^2 } \right),
\ee
\textbf{\citep{1952PASJ....4..100U, 1962MNRAS.125...21H}}.

The integral of the redistribution function over outgoing and incoming frequency are
\be
\int d\nu\ R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
& = & \frac{1}{4\pi} g(\vec{n},\vec{n}^\prime) \phi(\nu^\prime)
\ee 
and
\be
\int d\nu^\prime \ R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
& = & \frac{1}{4\pi} g(\vec{n},\vec{n}^\prime) \phi(\nu)
\ee 
where the right hand side is the usual Voigt function, the thermal average of the Lorentzian. The former result implies that the integrated source and sink terms for scattering cancel for $p=0$. In addition, $4\pi R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime)/\phi(\nu^\prime) $ is the normalized distribution for the outgoing $\vec{n}$ and $\nu$ given the incoming $\vec{n}^\prime$ and $\nu^\prime$.

This probability distribution can be used to define the moments of the frequency shift \citep{1962ApJ...135..195O}
\be
\langle \delta \nu^n \rangle & = & \frac{ \int d\nu^\prime (\nu^\prime-\nu)^n R}{\int d\nu^\prime R}
= \frac{1}{\phi(\nu)}
\int \frac{ d^3v}{ \pi^{3/2} v_{\rm th}^3} e^{-v^2/v_{\rm th}^2}\,
\left( \frac{\nu_0 \vec{v} \cdot (\vec{n}^\prime-\vec{n}) }{c} \right)^n
\left( \frac{\Gamma/4\pi^2}{ \left(\nu - \nu_0 - \nu_0 \vec{v} \cdot \vec{n}/c \right)^2 + (\Gamma/4\pi)^2 } \right),
\ee
which are functions of $\nu$, $\vec{n}$ and $\vec{n}^\prime$. These integrals can be evaluated in terms of the dimensionless moments of the parallel velocity distribution, defined as
\be
\langle u_\parallel^n \rangle(x,a) & = & \frac{a/\pi }{\mathcal{H}(x,a)} \int 
\frac{du_\parallel u_\parallel^n e^{-u_\parallel^2}  }{(x-u_\parallel)^2 + a^2}.
\ee
The end results for the first and second moments are
\be
\langle \delta \nu \rangle & = & -\langle u_\parallel \rangle \left( 1 - \vec{n} \cdot \vec{n}^\prime \right) \Delta 
\\
\langle \delta \nu^2 \rangle & = & 
\left[ \langle u_\parallel^2 \rangle
\left( 1 - \vec{n} \cdot \vec{n}^\prime \right)^2
+ \frac{1}{2} \left( 1 - \left( \vec{n} \cdot \vec{n}^\prime\right)^2 \right) \right] \Delta^2 .
\ee

For small frequency shifts $\nu^\prime-\nu$, the incoming intensity may be expanded as 
\be
I(\vec{x},\vec{n}^\prime,\nu^\prime) & \simeq  &
I(\vec{x},\vec{n}^\prime,\nu) 
+ 
\frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }(\nu^\prime-\nu)
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2} (\nu^\prime-\nu)^2 + ...
\ee 
and the Fokker-Planck expansion of $j_{\rm sc}$ is \citep{1994ApJ...427..603R}
\be
j_{\rm sc}(\vec{x},\vec{n},\nu) & \simeq  & 4\pi k \int d\Omega^\prime \int d\nu^\prime R(\vec{n},\nu; \vec{n}^\prime,\nu^\prime) 
\left[ 
I(\vec{x},\vec{n}^\prime,\nu) + \frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }(\nu^\prime-\nu)
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2 }(\nu^\prime-\nu)^2
\right]
\nonumber \\ 
& =& 
k\phi(\nu) \int d\Omega^\prime g 
\left[ 
I(\vec{x},\vec{n}^\prime,\nu) 
+ \frac{\partial I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu }
\langle \delta \nu \rangle
+ \frac{1}{2} \frac{\partial^2 I(\vec{x},\vec{n}^\prime,\nu)}{\partial \nu^2 }
\langle \delta \nu^2 \rangle
\right].
\ee 
To perform the angular integrals, the Eddington approximation for the angular dependence is inserted with the following result
\be
j_{\rm sc} & = & k\phi J - k\phi \Delta \langle u_\parallel \rangle \left( \frac{\partial J}{\partial \nu} - \frac{6}{5} \vec{n} \cdot \frac{\partial \vec{H}}{\partial \nu} \right)
+ \frac{1}{2} \Delta^2 k\phi \left[ 
\frac{\partial^2 J}{\partial \nu^2} \left( \frac{7}{5} \langle u_\parallel^2 \rangle + \frac{3}{10} \right)
- \frac{12}{5} \langle u_\parallel^2 \rangle 
\vec{n} \cdot \frac{\partial^2 \vec{H}}{\partial \nu^2} \right]
\nonumber \\ & \simeq & 
k\phi J - k\phi \Delta \langle u_\parallel \rangle  \frac{\partial J}{\partial \nu} 
+ \frac{1}{2} \Delta^2 k\phi \left( \frac{7}{5} \langle u_\parallel^2 \rangle + \frac{3}{10} \right)
\frac{\partial^2 J}{\partial \nu^2} .
\label{eq:jsc}
\ee
The first term in Equation (\ref{eq:jsc}), $k\phi J$, represents re-emission of the photon through de-excitation of the atom. It cancels the $-k\phi J$ term in Equation (\ref{eq:rteqn}) that corresponds to excitation of the atom. The terms involving frequency derivatives of $\vec{H}$, if carried through the calculation, end up giving terms smaller than the largest terms by a factor of $1/x^2$, which is small in the line wing. These terms are ignored from here onward.

@@

\right),

@@

Near the line center, there are two independent solutions which can be written
\be \label{eq:wkb_J}
J(n,\sigma,s) & = & a_1 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/2} J_{-3/4}\left( k_c \left[ \frac{\sigma}{\sigma_{\rm tp} }\right]^{2/3} \right)
\nonumber \\ &  + &  a_2 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/2} J_{3/4}\left( k_c \left[ \frac{\sigma}{\sigma_{\rm tp} }\right]^{2/3} \right),
\ee
where $J_\alpha(x)$ are Bessel functions of the first kind, $a_1$ and $a_2$ are normalization constants, $\sigma_{tp}$ is given in Equation (\ref{eq:sigma_tp}), and $k_c=3\kappa_n \Delta \sigma_{\rm tp}/2k$. The solution far from line center is 
\be \label{eq:wkb_J_far}
J(n,\sigma,s) & = & a_3 \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{1/6} Ai\left(k_c^{2/3} \left[ \left( \frac{\sigma}{\sigma_{\rm tp}} \right)^{2/3} - 1 \right]\right),
\ee
where $Ai(x)$ is the Airy function of the first kind. Again, $a_3$ is a normalization constant. We can enforce the discontinuity in Equation (\ref{eq:matching_condition_2}) to find 
\be \label{eq:a2}
a_2 & = & - \frac{\Gamma(7/4) \sigma_{\rm tp}}{ 2^{1/4} k_c^{3/4}} \frac{6^{1/2}}{8} n^2 \frac{E}{kR^3}.
\ee
The discontinuity sets the value of $a_2$, and the corresponding term in Equation (\ref{eq:wkb_J}) is non-resonant. The discontinuity does not set the value of $a_1$ as that solution does not have a term linear in $\sigma$. Matched asymptotic expansions \citep{1999amms.book.....B} allows us to relate the normalization coefficients $a_1$ and $a_3$. The matching conditions are
\be \label{eq:a1/a2}
\frac{a_1}{a_2} & = & \frac{ \sin \left( \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} + \frac{\pi}{8} \right) }{ \sin \left(  \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} - \frac{\pi}{8} \right) }
\ee
\be \label{eq:a3/a2}
\frac{a_3}{a_2} & = & \frac{ k_c^{-1/3} }{ \sin \left(  \frac{\kappa_n \sigma_{\rm tp} \Delta}{k} - \frac{\pi}{8} \right) }.
\ee
By setting the denominator of this expression equal to zero and solving for the eigenfrequency $\gamma_{nm}$ contained in $\sigma_{\rm tp}$, we find the dispersion relation in Equation (\ref{eq:gamma_nm}).
